---
title: "Cold stunned turtles take 2"
output: html_notebook
---



This document describes statistical analyses of cold stunned marine turtle data from Heather Harris. First load data and see what they look like.


```{r}

rm(list=ls())
source("cold_stun_functions.R")
library(tidyverse)

save.fig <- FALSE #TRUE

#The stranding data below were extracted from the stranding database using R script file in the strandings project. 

all.strandings <- read.csv(file = "data/WC_Strandings_Chelonia_Caretta_Lepidochelys_2019-06-13.csv")

# remove those with obvious human interactions, including fishery interactions
all.strandings %>% rownames_to_column() %>%
  filter(is.na(Fishery_Interaction)) %>% 
  filter(toupper(Human_Interaction) != "YES") %>% #-> tmp
  filter(!is.na(Latitude)) %>% #-> tmp
  filter(Latitude > 34.45 & Latitude < 51) %>% #-> tmp
  filter(Year_Initially_Observed >= 2003) %>%
  mutate(ID = rowname,
         Species = as.factor(ifelse(Species == "mydas", "CM",
                                    ifelse(Species == "caretta", "CC", "LO"))),
         CCL_cm = Curved_Carapace_Length,
         Weight_kg = Weight,
         Longitude = ifelse(Longitude < 0, Longitude, -1 * Longitude),
         Date = as.Date(paste(Year_Initially_Observed, 
                              Month_Initially_Observed, 
                              Day_Initially_Observed, sep = "-"))) %>%
  select(ID, Species, Date, State, Latitude, Longitude, CCL_cm, Weight_kg) -> all.strandings.1

dat0 <- read.csv(file = "data/ColdStun_data_June2019.csv", 
                 header = TRUE)  %>%
  rownames_to_column() %>%
  select(-c(Field_ID, Database_ID, NMFS_ID, Other_ID, Location, City, County, Country)) %>%
  mutate(Age = as.factor(toupper(Age)),
         Sex = as.factor(toupper(Sex)),
         Condition = as.factor(toupper(Condition)),
         ID = rowname,
         Species = as.factor(ifelse(Common_Name == "Green sea turtle", "CM",
                                    ifelse(Common_Name == "Loggerhead", "CC", "LO"))),
         Body_Temp_C = Confirmed.body.temp.C,
         Year = Year_Initially_Observed,
         Month = Month_Initially_Observed,
         Day = Day_Initially_Observed,
         Date = as.Date(paste(Year, Month, Day, sep = "-"))) %>%
  select(ID, Age, State, Latitude, Longitude, 
         Sex, Weight_kg, CCL_cm, CCW_cm,
         Body_Temp_C, Date, Species) %>%
  filter(Latitude > 34.45)

summary(dat0)

```


Removed those found south of Pt Conception (34.45N), as Deena did (true?).  

Then, remove those in dat0 from all.strandings.1.
```{r}

all.strandings.2 <- anti_join(all.strandings.1, dat0, by = c("Species", "Date", "Latitude")) 

```


Look at size measurements - compare with other stranding data.  May be good to show these are different from other stranded turtles. 


```{r}
p3 <- ggplot() + 
  geom_point(data = dat0,
             aes(x = CCL_cm, y = Weight_kg, color = Species, size = CCW_cm)) + 
  geom_point(data = all.strandings.2,
             aes(x = CCL_cm, y = Weight_kg, color = Species),
             shape = 1,
             size = 3)

if (save.fig)
  ggsave(plot = p3,
         device = "png",
         dpi = 600,
         filename = "figures/CCL_Mass_CCW.png")
p3
```

Looks like one loggerhead has incorrect weight... CCL and CCW for that individual is identical! 

```{r}
dat0 %>% filter(Species == "CC")
```

```{r}
CCL.dat0 <- select(dat0, CCL_cm, Species, Date) %>%
  mutate(data = "Cold") %>%
  na.omit()

CCL.other <- select(all.strandings.2, CCL_cm, Species, Date) %>% 
  mutate(data = "Other") %>%
  na.omit()

CCL.all <- rbind(CCL.dat0, CCL.other) %>%
  mutate(data = as.factor(data),
         Species = as.factor(Species))

CCL.all %>% group_by(data) %>% 
  summarise(mean = mean(CCL_cm),
            SE = sqrt(var(CCL_cm))/sqrt(n()))

```

So, there isn't any statistical difference between the two groups. An ANOVA confirms this - not included. 

```{r}
ggplot() + 
  geom_histogram(data = CCL.all,
                 aes(x = CCL_cm),
                 bins = 20) + 
  facet_grid(cols = vars(data))
```



Change in body size over time?
```{r}
p <- ggplot() + 
  geom_point(data = dat0,
             aes(x = Date, y = CCL_cm, color = Species, size = Latitude),
             alpha = 0.6) + 
  geom_point(data = all.strandings.2,
             aes(x = Date, y = CCL_cm, color = Species, size = Latitude),
             shape = 1,
             size = 3)

if (save.fig)
    ggsave(plot = p,
         device = "png",
         dpi = 600,
         filename = "figures/date_CCL_latitude.png")

p
```

Empty circles are from non-cold-stun animals. Not really for all turtles but I'm sure there is a positive relationship for loggerheads but it has a small sample size (n = 4). Greens may have a negative relationship but also has a small sample size (n = 5). 

Small sample sizes don't let me say too much about greens or loggerheads.  Olive ridleys may be getting a bit smaller? This was just looking at the cold-stun dataset... 

```{r}
dat0 %>% filter(Species == "LO") %>% 
  mutate(days = as.numeric(Date - min(Date))) -> dat0.LO

fit.LO.CCL <- lm(CCL_cm ~ days, data = dat0.LO)
summary(fit.LO.CCL)

```

When we pool both datasets, 

```{r}
CCL.all %>% filter(Species == "LO") %>% 
  mutate(days = as.numeric(Date - min(Date))) -> CCL.LO

fit.LO.CCL <- lm(CCL_cm ~ days, data = CCL.LO)
summary(fit.LO.CCL)

```

Not really... so, we don't think there was any change in CCL over time. 

The temperature data in the dataset are not very useful because some of them come from quite far away. We should probably get better data. Wind speed and direction prior to and at the day of stranding should be looked at. 

```{r, cache=TRUE, message=FALSE, warning=FALSE}
library(sp)
library(ncdf4)

# distances from each stranding within which SST data are extracted. It is measured from the 
# center (stranding location) +/- ds (in km). So, in effect SST is extracted from 2 * ds x 2 * ds
# area. Note they are not that different at the scale of 3 km to 10 km. 

ds <- c(3, 5, 10)
cold.stun.oce <- get_ocedata_fcn(ds, dat0, dataset = "coldstun")
all.stranding.oce<- get_ocedata_fcn(ds, all.strandings.2, dataset = "all")

```


Once data are downloaded, combine it with the original dataset.  CCL was not measured for 3 with body temperature... kinda odd? Only ridleys are available... 

```{r}
dat_01 <- cbind(dat0, 
                cold.stun.oce$sst01.df, 
                cold.stun.oce$sst01.min.df, 
                cold.stun.oce$sst01.anom.df, 
                cold.stun.oce$sst01.lag30d.df, 
                cold.stun.oce$sst01.min.lag30d.df, 
                cold.stun.oce$sst01.sd.df, 
                cold.stun.oce$sst01.anom.lag30d.df) %>%
  mutate(SST_1_CV = SST_1_SD/SST_1,
         SST_2_CV = SST_2_SD/SST_2,
         SST_3_CV = SST_3_SD/SST_3)

p4 <- ggplot(data = dat_01) + 
  geom_point(aes(x = SST_1_min, 
                 y = Body_Temp_C),
             size = 3) +
  geom_abline(slope = 1, intercept = 0) +
  xlim(c(8, 12)) +
  labs(y = "Body temperature (C)", 
       x = "Monthly composite SST minimum (3 km x 3 km, C)") 


if (save.fig)
  ggsave(plot = p4,
         device = "png",
         dpi = 600,
         filename = "figures/SST01_1_min_BodyTemp.png")

p4
```

There is one that had very low body temperature relative to the ambient SST.  SST_3_min is the monthly composite minimum of SST at 10 km by 10 km area. SST itself is measured at a 0.01 degree resolution. 

```{r}
dat0 %>% filter(Body_Temp_C < 8) 
```

It was an olive ridley that stranded in Oregon on 2018-12-13.  Anything specific about this turtle? For example, the body temp was measured hours after it was retrieved? 

Using 14-day composite doesn't change results much, nor the box around each stranding. See next 2 plots. 

```{r}
dat_0125 <- cbind(dat0, 
              cold.stun.oce$sst0125.df, 
              cold.stun.oce$sst0125.min.df, 
              cold.stun.oce$sst0125.lag30d.df, 
              cold.stun.oce$sst0125.min.lag30d.df, 
              cold.stun.oce$sst0125.sd.df)

p4 <- ggplot(data = filter(dat_0125, SST0125_3_min < Inf)) + 
  geom_point(aes(x = SST0125_3_min, 
                 y = Body_Temp_C), 
             size = 3) +
  geom_abline(slope = 1, intercept = 0) +
  xlim(c(9, 13))

if (save.fig)
  ggsave(plot = p4,
         device = "png",
         dpi = 600,
         filename = "figures/SST0125_3_min_BodyTemp.png")

p4
```

We get smaller sample size because of some missing data points for the minimum SST.  


```{r}

p4 <- ggplot(data = filter(dat_0125, SST0125_1_min < Inf)) + 
  geom_point(aes(x = SST0125_1_min, 
                 y = Body_Temp_C),
             size = 3) +
  geom_abline(slope = 1, intercept = 0) 
  #xlim(c(9, 13))

if (save.fig)
  ggsave(plot = p4,
         device = "png",
         dpi = 600,
         filename = "figures/SST0125_3_min_BodyTemp.png")

p4
```

Only 5 data points are available for the smaller spatial span... 


```{r}

p5 <- ggplot(data = dat_01) + 
  geom_point(aes(x = Date, 
                 y = SST_3_min, 
                 color = Species),
             size = 3) + 
  labs(x = "", y = "Monthly composite SST minimum (10 km x 10 km, C)") + 
  theme(legend.position = c(0.9, 0.8))

if (save.fig)
  ggsave(plot = p5,
         device = "png",
         dpi = 600,
         filename = "figures/date_SST01_3_min.png")

p5
```


```{r}

p5 <- ggplot(data = dat_01) + 
  geom_point(aes(x = Date, 
                 y = SST_2_min, 
                 color = Species),
             size = 3) + 
  labs(x = "", y = "Monthly composite SST minimum (5 km x 5 km, C)") + 
  theme(legend.position = c(0.9, 0.8))

if (save.fig)
  ggsave(plot = p5,
         device = "png",
         dpi = 600,
         filename = "figures/date_SST01_2_min.png")

p5
```


```{r}

p5 <- ggplot(data = dat_01) + 
  geom_point(aes(x = Date, 
                 y = SST_1_min, 
                 color = Species),
             size = 3) + 
  labs(x = "", y = "Monthly composite SST minimum (3 km x 3 km, C)") + 
  theme(legend.position = c(0.9, 0.8))

if (save.fig)
  ggsave(plot = p5,
         device = "png",
         dpi = 600,
         filename = "figures/date_SST01_1_min.png")

p5
```


```{r}

p5 <- ggplot(data = filter(dat_0125, SST0125_3_min < Inf)) + 
  geom_point(aes(x = Date, 
                 y = SST0125_3_min, 
                 color = Species),
             size = 3) + 
  labs(x = "", y = "14-day composite SST minimum (10 km x 10 km, C)") + 
  theme(legend.position = c(0.9, 0.8))

if (save.fig)
  ggsave(plot = p5,
         device = "png",
         dpi = 600,
         filename = "figures/date_SST0125_2_min.png")

p5
```


```{r}

p5 <- ggplot(data = filter(dat_0125, SST0125_2_min < Inf)) + 
  geom_point(aes(x = Date, 
                 y = SST0125_2_min, 
                 color = Species),
             size = 3) + 
  labs(x = "", y = "14-day composite SST minimum (5 km x 5 km, C)") + 
  theme(legend.position = c(0.9, 0.8))

if (save.fig)
  ggsave(plot = p5,
         device = "png",
         dpi = 600,
         filename = "figures/date_SST0125_2_min.png")

p5
```



```{r}

p5 <- ggplot(data = filter(dat_0125, SST0125_1_min < Inf)) + 
  geom_point(aes(x = Date, 
                 y = SST0125_1_min, 
                 color = Species),
             size = 3) + 
  labs(x = "", y = "14-day composite SST minimum (3 km x 3 km, C)") + 
  theme(legend.position = c(0.9, 0.8))

if (save.fig)
  ggsave(plot = p5,
         device = "png",
         dpi = 600,
         filename = "figures/date_SST0125_1_min.png")

p5
```


```{r}
p6 <- ggplot(data = dat_01) + 
  geom_point(aes(x = Date, 
                 y = SST_3_ANOM, 
                 color = Species),
             size = 3)

if (save.fig)
  ggsave(plot = p6,
         device = "png",
         dpi = 600,
         filename = "figures/date_SST01_3_anom.png")

p6
```


```{r}
p7 <- ggplot(data = dat_01) + 
  geom_point(aes(x = Date, 
                 y = SST_3_CV, 
                 color = Species),
             size = 3)

if (save.fig)
  ggsave(plot = p7,
         device = "png",
         dpi = 600,
         filename = "figures/date_SST01_3_cv.png")

p7
```


```{r}
p9 <- ggplot(data = dat_01) + 
  geom_point(aes(x = Date, 
                 y = SST_3_min_lag30, 
                 color = Species),
             size = 3)

if (save.fig)
  ggsave(plot = p9,
         device = "png",
         dpi = 600,
         filename = "figures/date_SST01_3_min_lag30.png")

p9

```



```{r}
p8 <- ggplot(data = dat_01) + 
  geom_point(aes(x = Date, 
                 y = SST_3_ANOM_lag30, 
                 color = Species),
             size = 3)

if (save.fig)
  ggsave(plot = p8,
         device = "png",
         dpi = 600,
         filename = "figures/date_SST01_3_anom_lag30.png")

p8
```


```{r}
p10 <- ggplot(data = dat_01) + 
  geom_point(aes(x = Date, 
                 y = SST_2_min_lag30, 
                 color = Species),
             size = 3)


if (save.fig)
  ggsave(plot = p10,
         device = "png",
         dpi = 600,
         filename = "figures/date_SST01_2_min_lag30.png")


p10
```



```{r}
p11 <- ggplot(data = dat_01) + 
  geom_point(aes(x = Date, 
                 y = SST_2_ANOM_lag30, 
                 color = Species),
             size = 3)

if (save.fig)
  ggsave(plot = p11,
         device = "png",
         dpi = 600,
         filename = "figures/date_SST01_2_anom_lag30.png")

p11
```