---
title: "Cold stunned marine turtles"
output: html_notebook
---

This document describes statistical analyses of cold stunned marine turtle data from Heather Harris. First load data and see what they look like.


```{r}

rm(list=ls())
source("cold_stun_functions.R")
library(tidyverse)

save.fig <- FALSE #TRUE

dat0 <- read.csv(file = "data/ColdStun_data_June2019.csv", 
                 header = TRUE)  %>%
  rownames_to_column() %>%
  select(-c(Field_ID, Database_ID, NMFS_ID, Other_ID, Location, City, County, Country)) %>%
  mutate(Age = as.factor(toupper(Age)),
         Sex = as.factor(toupper(Sex)),
         Condition = as.factor(toupper(Condition)),
         ID = rowname,
         Species = as.factor(ifelse(Common_Name == "Green sea turtle", "CM",
                                    ifelse(Common_Name == "Loggerhead", "CC", "LO"))),
         Body_Temp_C = Confirmed.body.temp.C,
         Year = Year_Initially_Observed,
         Month = Month_Initially_Observed,
         Day = Day_Initially_Observed,
         Date = as.Date(paste(Year, Month, Day, sep = "-"))) %>%
  select(ID, Age, State, Latitude, Longitude, 
         Sex, Weight_kg, CCL_cm, CCW_cm,
         Body_Temp_C, Date, Species) %>%
  filter(Latitude > 34.45)

summary(dat0)

```


Removed those found south of Pt Conception (34.45N), as Deena did (true?).  

Look at size measurements:

```{r}
p2 <- ggplot(data = dat0) +
  geom_point(aes(x = CCL_cm, y = CCW_cm, color = Species, size = Weight_kg))

if (save.fig)
  ggsave(plot = p2,
         device = "png",
         dpi = 600,
         filename = "figures/CCL_CCW_Mass.png")

p2
```


```{r}
p3 <- ggplot(data = dat0) + 
  geom_point(aes(x = CCL_cm, y = Weight_kg, color = Species, size = CCW_cm))

if (save.fig)
  ggsave(plot = p3,
         device = "png",
         dpi = 600,
         filename = "figures/CCL_Mass_CCW.png")
p3
```

Looks like one loggerhead has incorrect weight... CCL and CCW for that individual is identical! 

```{r}
dat0 %>% filter(Species == "CC")
```

The temperature data in the dataset are not very useful because some of them come from quite far away. We should probably get better data. Wind speed and direction prior to and at the day of stranding should be looked at. 

```{r, cache=TRUE, message=FALSE}
library(sp)
library(ncdf4)

# distances from each stranding within which SST data are extracted. It is measured from the 
# center (stranding location) +/- ds (in km). So, in effect SST is extracted from 2 * ds x 2 * ds
# area. Note they are not that different at the scale of 1 km to 10 km. 

ds <- c(3, 5, 10)
d.n <- length(ds)   # this should match the number of dx's below. 

# Get SST for 1, 5, 10 km around each point. 
latlon <- select(dat0, Latitude, Longitude) %>% 
  transmute(X = Longitude, Y = Latitude,
            BeginX_1 = NA, BeginY_1 = NA, EndX_1 = NA, EndY_1 = NA,
            BeginX_2 = NA, BeginY_2 = NA, EndX_2 = NA, EndY_2 = NA,
            BeginX_3 = NA, BeginY_3 = NA, EndX_3 = NA, EndY_3 = NA) 

UTM <- data.frame(X = NA, Y = NA,
                  BeginX_1 = NA, BeginY_1 = NA, EndX_1 = NA, EndY_1 = NA,
                  BeginX_2 = NA, BeginY_2 = NA, EndX_2 = NA, EndY_2 = NA,
                  BeginX_3 = NA, BeginY_3 = NA, EndX_3 = NA, EndY_3 = NA,
                  zone = NA)

# because the range of latitude is quite large, approximate center is moved along each data point.
# convert all lat/lon pairs to UTM coordinates in the appropriate zone. 
centers.UTM <- vector(mode = "list", length = nrow(latlon))
k <- 1
for (k in 1:nrow(dat0)){
  approx.center <- data.frame(X = -120, Y = dat0[k, "Latitude"])
  sp::coordinates(approx.center) <- c("X", "Y")
  sp::proj4string(approx.center) <- CRS("+proj=longlat +datum=WGS84")
  
  # North of ~ Pt Conception is zone 10, and south of it is zone 11
  zone <- ifelse(dat0[k, "Latitude"] > 34.46, 10, 11)
  
  centers.UTM[[k]] <- spTransform(approx.center,
                            CRS(paste0("+proj=utm +zone=", zone, " ellps=WGS84")))
  
  tmp <- latlon2sp(latlon[k, c("X", "Y")], 
                   center.UTM = centers.UTM[[k]], zone )
  
  UTM[k, ] <- c(tmp@data$newX, tmp@data$newY,
                tmp@data$newX - ds[1], tmp@data$newY - ds[1], 
                tmp@data$newX + ds[1], tmp@data$newY + ds[1],
                tmp@data$newX - ds[2], tmp@data$newY - ds[2], 
                tmp@data$newX + ds[2], tmp@data$newY + ds[2],
                tmp@data$newX - ds[3], tmp@data$newY - ds[3], 
                tmp@data$newX + ds[3], tmp@data$newY + ds[3],
                zone)
  
}

# This loop should go over all dx's (2 each) - Begin and End (X,Y)
k <- k1 <- 1
for (k in 1:(2*d.n)){
  tmp <- data.frame(newX = UTM[, (2*k+1)],
                    newY = UTM[, (2*k+2)])
  for (k1 in 1:nrow(tmp)){
    latlon[k1, (2*k+1):(2*k+2)] <- sp2latlon(tmp[k1,], 
                                             center.UTM = centers.UTM[[k1]], 
                                             zone = UTM[k1, "zone"])@coords[, c("X", "Y")]
    
  }
  
}

latlon$Date <- dat0$Date

sst <- sst_anom <- sst_sd <- sst_lag30d <- sst_anom_30d <- matrix(nrow = nrow(latlon),
                                                                  ncol = d.n)

sst0125 <- sst0125_min <- sst0125_sd <-  matrix(nrow = nrow(latlon),
                                                ncol = d.n)

sst_min <- sst_min_lag30d <- air_temp <- air_temp_30d <- matrix(nrow = nrow(latlon),
                                                                ncol = d.n)
sst0125_lag30d <- sst0125_min_lag30d <- sst0125_sd_lag30d <- matrix(nrow = nrow(latlon),
                                               ncol = d.n)

wind <- wind_max <- wind_30d <- wind_max_30d <- matrix(nrow = nrow(latlon),
                                                       ncol = 1)
k <- k1 <- 1

# for all rows
for (k in 1:nrow(sst)){
  # wind comes in 0.25 degree resolutions so we just pick one without worrying about other spatial
  # resolutions
  out.file.name.wind <- paste0("data/ncfiles/turtle_", dat0[k, "ID"], "_",
                               "wind.nc")
  
  out.file.name.wind.30d <- paste0("data/ncfiles/turtle_", dat0[k, "ID"], "_",
                                   "wind_30d.nc")
  
  # for number of distances
  for (k1 in 1:d.n){
    out.file.name <- paste0("data/ncfiles/turtle_", dat0[k, "ID"], "_",
                                             ds[k1], "km_sst.nc")
    out.file.name.anom <- paste0("data/ncfiles/turtle_", dat0[k, "ID"], "_",
                                             ds[k1], "km_sstanom.nc")
    out.file.name.lag30d<- paste0("data/ncfiles/turtle_", dat0[k, "ID"], "_",
                                             ds[k1], "km_sst_lag30d.nc")
    out.file.name.anom.lag30d <- paste0("data/ncfiles/turtle_", dat0[k, "ID"], "_",
                                             ds[k1], "km_sstanom_lag30d.nc")
    out.file.name.air.temp<- paste0("data/ncfiles/turtle_", dat0[k, "ID"], "_",
                                             ds[k1], "km_airtemp.nc")
    out.file.name.air.temp.lag30d <- paste0("data/ncfiles/turtle_", dat0[k, "ID"], "_",
                                             ds[k1], "km_airtemp_lag30d.nc")
    out.file.name.sst.0125 <- paste0("data/ncfiles/turtle_", dat0[k, "ID"], "_",
                                             ds[k1], "km_sst0125.nc")
    out.file.name.sst.0125.lag30d <- paste0("data/ncfiles/turtle_", dat0[k, "ID"], "_",
                                             ds[k1], "km_sst0125_lag30d.nc")
    
    
    xlim <- c(latlon[k, (4*k1 - 1)],
              latlon[k, (4*k1 + 1)])
    ylim <- c(latlon[k, (4*k1)],
              latlon[k, (4*k1 + 2)])
    #tlim <- c(latlon[k, "Date"] - 7,
    #          latlon[k, "Date"] + 7)
  
    datafileID <- nc_open(out.file.name.sst.0125)
    lon <- ncvar_get(datafileID, varid="longitude")
    lat <- ncvar_get(datafileID, varid="latitude")
    time <- ncvar_get(datafileID, varid="time")
    time <- as.POSIXlt(time, origin='1970-01-01',tz= "GMT")  
    mat <- ncvar_get(datafileID, varid = 'sst')
    nc_close(datafileID)
    
    sst0125[k, k1] <- mean(mat, na.rm = T)
    sst0125_sd[k, k1] <- sqrt(var(as.vector(mat), na.rm = T))
    sst0125_min[k, k1] <- min(mat, na.rm = T)
      
    datafileID <- nc_open(out.file.name.sst.0125.lag30d)
    lon <- ncvar_get(datafileID, varid="longitude")
    lat <- ncvar_get(datafileID, varid="latitude")
    time <- ncvar_get(datafileID, varid="time")
    time <- as.POSIXlt(time, origin='1970-01-01',tz= "GMT")  
    mat <- ncvar_get(datafileID, varid = 'sst')
    nc_close(datafileID)
    
    sst0125_lag30d[k, k1] <- mean(mat, na.rm = T)
    sst0125_sd_lag30d[k, k1] <- sqrt(var(as.vector(mat), na.rm = T))
    sst0125_min_lag30d[k, k1] <- min(mat, na.rm = T)
    
    datafileID <- nc_open(out.file.name)
    lon <- ncvar_get(datafileID, varid="longitude")
    lat <- ncvar_get(datafileID, varid="latitude")
    time <- ncvar_get(datafileID, varid="time")
    time <- as.POSIXlt(time, origin='1970-01-01',tz= "GMT")  
    mat <- ncvar_get(datafileID, varid = 'sst')
    nc_close(datafileID)
    
    sst[k, k1] <- mean(mat, na.rm = T)
    sst_sd[k, k1] <- sqrt(var(as.vector(mat), na.rm = T))
    sst_min[k, k1] <- min(mat, na.rm = T)
    
    datafileID <- nc_open(out.file.name.anom)
    lon <- ncvar_get(datafileID, varid="longitude")
    lat <- ncvar_get(datafileID, varid="latitude")
    time <- ncvar_get(datafileID, varid="time")
    time <- as.POSIXlt(time, origin='1970-01-01',tz= "GMT")  # true? check with nc files (5/30/2019)
    mat <- ncvar_get(datafileID, varid = 'sstAnom')
    nc_close(datafileID)
    
    sst_anom[k, k1] <- mean(mat, na.rm = T)
    
    datafileID <- nc_open(out.file.name.lag30d)
    lon <- ncvar_get(datafileID, varid="longitude")
    lat <- ncvar_get(datafileID, varid="latitude")
    time <- ncvar_get(datafileID, varid="time")
    time <- as.POSIXlt(time, origin='1970-01-01',tz= "GMT")  
    mat <- ncvar_get(datafileID, varid = 'sst')
    nc_close(datafileID)
    
    sst_lag30d[k, k1] <- mean(mat, na.rm = T)
    sst_min_lag30d[k, k1] <- min(mat, na.rm = T)
  
    datafileID <- nc_open(out.file.name.anom.lag30d)
    lon <- ncvar_get(datafileID, varid="longitude")
    lat <- ncvar_get(datafileID, varid="latitude")
    time <- ncvar_get(datafileID, varid="time")
    time <- as.POSIXlt(time, origin='1970-01-01',tz= "GMT")  
    mat <- ncvar_get(datafileID, varid = 'sstAnom')
    nc_close(datafileID)
    
    sst_anom_30d[k, k1] <- mean(mat, na.rm = T)
    
    
  }

  if (latlon[k, "Date"] < as.Date("2018-10-10")){
    datafileID <- nc_open(out.file.name.wind)
    lon <- ncvar_get(datafileID, varid="longitude")
    lat <- ncvar_get(datafileID, varid="latitude")
    time <- ncvar_get(datafileID, varid="time")
    time <- as.POSIXlt(time, origin='1970-01-01', tz= "GMT")  
    u <- ncvar_get(datafileID, varid = "u")
    v <- ncvar_get(datafileID, varid = "v")
    w <- ncvar_get(datafileID, varid = "w")
    
    nc_close(datafileID)
  } else {
    U <- v <- w <- NA
  }
  wind[k, 1] <- mean(w, na.rm = T)
  wind_max[k, 1] <- max(w, na.rm = T)
  #wind_u[k, 1] <- mean(u, na.rm = T)
  #wind_v[k, 1] <- mean(v, na.rm = T)
  
  if (((latlon[k, "Date"] > as.Date("2011-09-30")) & 
       (latlon[k, "Date"] - 30) < as.Date("2011-10-01"))){
    
    datafileID <- nc_open(out.file.name.wind.30d)
    datafileID2 <- nc_open(paste0(unlist(out.file.name.wind.30d, "\\.nc"), "_dat1.nc"))
    
    lon <- ncvar_get(datafileID, varid="longitude")
    lat <- ncvar_get(datafileID, varid="latitude")
    time <- ncvar_get(datafileID, varid="time")
    time <- as.POSIXlt(time, origin='1970-01-01', tz= "GMT")  
    u <- ncvar_get(datafileID, varid = "u")
    v <- ncvar_get(datafileID, varid = "v")
    w <- ncvar_get(datafileID, varid = "w")
    
    lon2 <- ncvar_get(datafileID2, varid="longitude")
    lat2 <- ncvar_get(datafileID2, varid="latitude")
    time2 <- ncvar_get(datafileID2, varid="time")
    time2 <- as.POSIXlt(time2, origin='1970-01-01', tz= "GMT")  
    u2 <- ncvar_get(datafileID2, varid = "u")
    v2 <- ncvar_get(datafileID2, varid = "v")
    w2 <- ncvar_get(datafileID2, varid = "w")
    
    nc_close(datafileID)
    nc_close(datafileID2)
    
    u <- abind::abind(u, u2)
    v <- abind::abind(v, v2)
    w <- abind::abind(w, w2)
    
  } else {
    if (latlon[k, "Date"] < as.Date("2018-10-10")){
      datafileID <- nc_open(out.file.name.wind.30d)
      lon <- ncvar_get(datafileID, varid="longitude")
      lat <- ncvar_get(datafileID, varid="latitude")
      time <- ncvar_get(datafileID, varid="time")
      time <- as.POSIXlt(time, origin='1970-01-01', tz= "GMT")  
      u <- ncvar_get(datafileID, varid = "u")
      v <- ncvar_get(datafileID, varid = "v")
      w <- ncvar_get(datafileID, varid = "w")
      
      nc_close(datafileID)
    } else {
      w <- u <- v <- NA
    }
  }
  
  wind_30d[k, 1] <- mean(w, na.rm = T)
  wind_max_30d[k, 1] <- max(w, na.rm = T)
  #wind_u_30d[k, 1] <- mean(u, na.rm = T)
  #wind_v_30d[k, 1] <- mean(v, na.rm = T)
}


```


Change in body size over time?
```{r}
p <- ggplot(data = dat0) + 
  geom_point(aes(x = Date, y = CCL_cm, color = Species, size = Latitude))

if (save.fig)
    ggsave(plot = p,
         device = "png",
         dpi = 600,
         filename = "figures/date_CCL_latitude.png")

p
```

Small sample sizes don't let me say too much about greens or loggerheads.  Olive ridleys may be getting a bit smaller? 

```{r}
dat0 %>% filter(Species == "LO") %>% 
  mutate(days = as.numeric(Date - min(Date))) -> dat0.LO

fit.LO.CCL <- lm(CCL_cm ~ days, data = dat0.LO)
summary(fit.LO.CCL)

```

NOt really... There is a "significant" negative relationship but not sure if this is just a result of randomness with the small sample size... 

Next, I look at effects of oceanographic variables on the number of strandings per year.  Note that this time definition is rather arbitrary (per year).  I probably should look into the number of strandings per oceanographic "season."  I'll do that later... for now, just year as a factor. 

Once data are downloaded, combine it with the original dataset.  CCL was not measured for 3 with body temperature... kinda odd? Only ridleys are available... 

```{r}
sst.df <- data.frame(sst)
colnames(sst.df) <- c("SST_1", "SST_2", "SST_3")

sst.anom.df <- data.frame(sst_anom)
colnames(sst.anom.df) <- c("SST_1_ANOM", "SST_2_ANOM", "SST_3_ANOM")

sst.sd.df <- data.frame(sst_sd)
colnames(sst.sd.df) <- c("SST_1_SD", "SST_2_SD", "SST_3_SD")

sst.min.df <- data.frame(sst_min)
colnames(sst.min.df) <- c("SST_1_min", "SST_2_min", "SST_3_min")

sst.lag30d.df <- data.frame(sst_lag30d)
colnames(sst.lag30d.df) <- c("SST_1_lag30", "SST_2_lag30", "SST_3_lag30")

sst.min.lag30d.df <- data.frame(sst_min_lag30d)
colnames(sst.min.lag30d.df) <- c("SST_1_min_lag30", "SST_2_min_lag30", "SST_3_min_lag30")

sst.anom.lag30d.df <- data.frame(sst_anom_30d)
colnames(sst.anom.lag30d.df) <- c("SST_1_ANOM_lag30", "SST_2_ANOM_lag30", "SST_3_ANOM_lag30")

sst0125.df <- data.frame(sst0125)
colnames(sst0125.df) <- c("SST0125_1", "SST0125_2", "SST0125_3")

sst0125.sd.df <- data.frame(sst0125_sd)
colnames(sst0125.sd.df) <- c("SST0125_1_SD", "SST0125_2_SD", "SST0125_3_SD")

sst0125.min.df <- data.frame(sst0125_min)
colnames(sst0125.min.df) <- c("SST0125_1_min", "SST0125_2_min", "SST0125_3_min")

sst0125.lag30d.df <- data.frame(sst0125_lag30d)
colnames(sst0125.lag30d.df) <- c("SST0125_1_lag30d", "SST0125_2_lag30d", "SST0125_3_lag30d")

# sst0125.lag30d.sd.df <- data.frame(sst0125_sd_lag30d)
# colnames(sst0125.lag30d.sd.df) <- c("SST0125_1_SD", "SST0125_2_SD", "SST0125_3_SD")

sst0125.lag30d.min.df <- data.frame(sst0125_min_lag30d)
colnames(sst0125.lag30d.df) <- c("SST0125_1_min_lag30", "SST0125_2_min_lag30", "SST0125_3_min_lag30")

wind.df <- data.frame(wind)

dat1 <- cbind(dat0, 
              sst.df, 
              sst.min.df, 
              sst.anom.df, 
              sst.lag30d.df, 
              sst.min.lag30d.df, 
              sst.sd.df, 
              sst.anom.lag30d.df,
              sst0125.df) %>%
  mutate(SST_1_CV = SST_1_SD/SST_1,
         SST_2_CV = SST_2_SD/SST_2,
         SST_3_CV = SST_3_SD/SST_3)

p4 <- ggplot(data = dat1) + 
  geom_point(aes(x = SST_3_min, 
                 y = Body_Temp_C, 
                 size = CCL_cm)) +
  xlim(c(9, 12))

if (save.fig)
  ggsave(plot = p4,
         device = "png",
         dpi = 600,
         filename = "figures/SST1_min_BodyTemp.png")

p4
```

```{r}

p5 <- ggplot(data = dat1) + 
  geom_point(aes(x = Date, 
                 y = SST_3_min, 
                 color = Species),
             size = 3)

if (save.fig)
  ggsave(plot = p5,
         device = "png",
         dpi = 600,
         filename = "figures/date_SST3_min.png")

p5
```

```{r}
p6 <- ggplot(data = dat1) + 
  geom_point(aes(x = Date, 
                 y = SST_3_ANOM, 
                 color = Species),
             size = 3)

if (save.fig)
  ggsave(plot = p6,
         device = "png",
         dpi = 600,
         filename = "figures/date_SST3anom.png")

p6
```


```{r}
p7 <- ggplot(data = dat1) + 
  geom_point(aes(x = Date, 
                 y = SST_3_CV, 
                 color = Species),
             size = 3)

if (save.fig)
  ggsave(plot = p7,
         device = "png",
         dpi = 600,
         filename = "figures/date_SST3cv.png")

p7
```


```{r}
p9 <- ggplot(data = dat1) + 
  geom_point(aes(x = Date, 
                 y = SST_3_min_lag30, 
                 color = Species),
             size = 3)

if (save.fig)
  ggsave(plot = p9,
         device = "png",
         dpi = 600,
         filename = "figures/date_SST3_min_lag30.png")

p9

```



```{r}
p8 <- ggplot(data = dat1) + 
  geom_point(aes(x = Date, 
                 y = SST_3_ANOM_lag30, 
                 color = Species),
             size = 3)

if (save.fig)
  ggsave(plot = p8,
         device = "png",
         dpi = 600,
         filename = "figures/date_SST3anom_lag30.png")

p8
```


```{r}
p10 <- ggplot(data = dat1) + 
  geom_point(aes(x = Date, 
                 y = SST_2_min_lag30, 
                 color = Species),
             size = 3)


if (save.fig)
  ggsave(plot = p10,
         device = "png",
         dpi = 600,
         filename = "figures/date_SST2_min_lag30.png")


p10
```



```{r}
p11 <- ggplot(data = dat1) + 
  geom_point(aes(x = Date, 
                 y = SST_2_ANOM_lag30, 
                 color = Species),
             size = 3)

if (save.fig)
  ggsave(plot = p11,
         device = "png",
         dpi = 600,
         filename = "figures/date_SST2anom_lag30.png")

p11
```

It seems like 2015 was a quite different year from all others - the reasons for these observed strandings for 2015 might have been different from others. So, to incorporate the climate effect, I downloaded El Nino index from here (https://www.esrl.noaa.gov/psd/enso/mei/). The data file consists of two-month values for each year, starting in 1979.  The current version (as of June 5, 2019) includes 2019 March/April.  Raw data file (meiv2_20190605.txt) was manipulated to remove lines that describe the datafile, rather than providing actual index values.  The new file is named meiv2_notext_20190605.txt. Missing values for the remainder of 2019 are indicated by -999.00. 

```{r}
library(data.table)
MEI.v2 <- fread(file = "data/meiv2_notext_20190605.txt")
colnames(MEI.v2) <- c("Year", "DJ", "JF", "FM", "MA", "AM", 
                      "MJ", "JJ", "JA", "AS", "SO", "ON", "ND") 
MEI.v2[MEI.v2==-999] <- NA

MEI.v2.long <- melt(MEI.v2, 
                    id.vars = "Year", 
                    value.name = "MEI.v2",
                    variable.name = "period")

month <- vector(mode = "numeric", length = nrow(MEI.v2.long))

month[MEI.v2.long$period == "DJ"] <- 1
month[MEI.v2.long$period == "JF"] <- 2
month[MEI.v2.long$period == "FM"] <- 3
month[MEI.v2.long$period == "MA"] <- 4
month[MEI.v2.long$period == "AM"] <- 5
month[MEI.v2.long$period == "MJ"] <- 6
month[MEI.v2.long$period == "JJ"] <- 7
month[MEI.v2.long$period == "JA"] <- 8
month[MEI.v2.long$period == "AS"] <- 9
month[MEI.v2.long$period == "SO"] <- 10
month[MEI.v2.long$period == "ON"] <- 11
month[MEI.v2.long$period == "ND"] <- 12

MEI.v2.long$month <- month
MEI.v2.long$year.month.num <- MEI.v2.long$Year + MEI.v2.long$month/12

MEI.v2.long <- arrange(MEI.v2.long, Year) %>% 
  mutate(year.month = paste(Year, month, sep = "."))

MEI.v2.long %>% 
  group_by(Year) %>%
  summarize(mean.MEI = mean(MEI.v2),
            SD.MEI = sqrt(var(MEI.v2, na.rm = T)),
            min.MEI = min(MEI.v2, na.rm = T),
            max.MEI = max(MEI.v2, na.rm = T)) -> MEI.v2.byYr

MEI.v2.lag1yr.long <- data.frame(Year = MEI.v2.long$Year[13:nrow(MEI.v2.long)],
                                 month = MEI.v2.long$month[13:nrow(MEI.v2.long)],
                                 MEI.v2.lag1yr = MEI.v2.long$MEI.v2[1:(nrow(MEI.v2.long)-12)]) %>% 
  mutate(year.month = paste(Year, month, sep = "."))

MEI.v2.lag1yr.byYr <- MEI.v2.lag1yr.long %>%
  group_by(Year) %>%
  summarize(mean.MEI.lag1yr = mean(MEI.v2.lag1yr),
            SD.MEI.lag1yr = sqrt(var(MEI.v2.lag1yr, na.rm = T)),
            min.MEI.lag1yr = min(MEI.v2.lag1yr, na.rm = T),
            max.MEI.lag1yr = max(MEI.v2.lag1yr, na.rm = T))


MEI.v2.lag6mo.long <- data.frame(Year = MEI.v2.long$Year[7:nrow(MEI.v2.long)],
                                 month = MEI.v2.long$month[7:nrow(MEI.v2.long)],
                                 MEI.v2.lag6mo = MEI.v2.long$MEI.v2[1:(nrow(MEI.v2.long)-6)]) %>% 
  mutate(year.month = paste(Year, month, sep = "."))

MEI.v2.lag6mo.long %>% 
  group_by(Year) %>%
  summarize(mean.MEI.lag6mo = mean(MEI.v2.lag6mo),
            SD.MEI.lag6mo = sqrt(var(MEI.v2.lag6mo, na.rm = T)),
            min.MEI.lag6mo = min(MEI.v2.lag6mo, na.rm = T),
            max.MEI.lag6mo = max(MEI.v2.lag6mo, na.rm = T)) -> MEI.v2.lag6mo.byYr


```

Pull in PDO
```{r}
PDO <-  fread(file = "data/PDO_notext_20190605.txt")
PDO.long <- melt(PDO, 
                 id.vars = "YEAR", 
                 value.name = "PDO",
                 variable.name = "Month")

month <- vector(mode = "numeric", length = nrow(PDO.long))

month[PDO.long$Month == "JAN"] <- 1
month[PDO.long$Month == "FEB"] <- 2
month[PDO.long$Month == "MAR"] <- 3
month[PDO.long$Month == "APR"] <- 4
month[PDO.long$Month == "MAY"] <- 5
month[PDO.long$Month == "JUN"] <- 6
month[PDO.long$Month == "JUL"] <- 7
month[PDO.long$Month == "AUG"] <- 8
month[PDO.long$Month == "SEP"] <- 9
month[PDO.long$Month == "OCT"] <- 10
month[PDO.long$Month == "NOV"] <- 11
month[PDO.long$Month == "DEC"] <- 12

PDO.long$month <- month

arrange(PDO.long, YEAR) %>% 
  mutate(year.month = paste(YEAR, month, sep = "."),
         Year = YEAR) -> PDO.long

PDO.long %>% 
  group_by(Year) %>%
  summarize(mean.PDO = mean(PDO, na.rm = T),
            SD.PDO = sqrt(var(PDO, na.rm = T)),
            min.PDO = min(PDO, na.rm = T),
            max.PDO = max(PDO, na.rm = T)) -> PDO.byYr

PDO.lag1yr <- data.frame(Year = PDO.long$Year[13:nrow(PDO.long)],
                         month = PDO.long$month[13:nrow(PDO.long)],
                         PDO.lag1yr = PDO.long$PDO[1:(nrow(PDO.long)-12)]) %>% 
  mutate(year.month = paste(Year, month, sep = "."))

PDO.lag6mo <- data.frame(Year = PDO.long$Year[7:nrow(PDO.long)],
                         month = PDO.long$month[7:nrow(PDO.long)],
                         PDO.lag6mo = PDO.long$PDO[1:(nrow(PDO.long)-6)]) %>% 
  mutate(year.month = paste(Year, month, sep = "."))


PDO.lag1yr.byYr <- PDO.lag1yr %>%
  group_by(Year) %>%
  summarize(mean.PDO.lag1yr = mean(PDO.lag1yr),
            SD.PDO.lag1yr = sqrt(var(PDO.lag1yr, na.rm = T)),
            min.PDO.lag1yr = min(PDO.lag1yr, na.rm = T),
            max.PDO.lag1yr = max(PDO.lag1yr, na.rm = T))

PDO.lag6mo.byYr <- PDO.lag6mo %>%
  group_by(Year) %>%
  summarize(mean.PDO.lag6mo = mean(PDO.lag6mo),
            SD.PDO.lag6mo = sqrt(var(PDO.lag6mo, na.rm = T)),
            min.PDO.lag6mo = min(PDO.lag6mo, na.rm = T),
            max.PDO.lag6mo = max(PDO.lag6mo, na.rm = T))

```

ONI:
```{r}
# data are from here - there are several versions on the Internet and they are slightly different...
# https://origin.cpc.ncep.noaa.gov/products/analysis_monitoring/ensostuff/detrend.nino34.ascii.txt
ONI <- fread(file = "data/ONI_20190605.txt")

arrange(ONI, YR) %>%
  mutate(year.month = paste(YR, MON, sep = "."),
         Year = YR) -> ONI.long

ONI.long %>%
  group_by(Year) %>%
  summarise(mean.ONI = mean(ANOM, na.rm = T),
            SD.ONI = sqrt(var(ANOM, na.rm = T)),
            min.ONI = min(ANOM, na.rm = T),
            max.ONI = max(ANOM, na.rm = T)) -> ONI.byYr

ONI.lag1yr <- data.frame(Year = ONI.long$Year[13:nrow(ONI.long)],
                         month = ONI.long$MON[13:nrow(ONI.long)],
                         ONI.lag1yr = ONI.long$ANOM[1:(nrow(ONI.long)-12)]) %>% 
  mutate(year.month = paste(Year, month, sep = "."))

ONI.lag6mo <- data.frame(Year = ONI.long$Year[7:nrow(ONI.long)],
                         month = ONI.long$MON[7:nrow(ONI.long)],
                         ONI.lag6mo = ONI.long$ANOM[1:(nrow(ONI.long)-6)]) %>% 
  mutate(year.month = paste(Year, month, sep = "."))

ONI.lag1yr.byYr <- ONI.lag1yr %>%
  group_by(Year) %>%
  summarize(mean.ONI.lag1yr = mean(ONI.lag1yr),
            SD.ONI.lag1yr = sqrt(var(ONI.lag1yr, na.rm = T)),
            min.ONI.lag1yr = min(ONI.lag1yr, na.rm = T),
            max.ONI.lag1yr = max(ONI.lag1yr, na.rm = T))

ONI.lag6mo.byYr <- ONI.lag6mo %>%
  group_by(Year) %>%
  summarize(mean.ONI.lag6mo = mean(ONI.lag6mo),
            SD.ONI.lag6mo = sqrt(var(ONI.lag6mo, na.rm = T)),
            min.ONI.lag6mo = min(ONI.lag6mo, na.rm = T),
            max.ONI.lag6mo = max(ONI.lag6mo, na.rm = T))

```



Combine together

```{r}
dat1 %>% mutate(Year = lubridate::year(Date),
                year.month = paste(lubridate::year(Date),
                                   lubridate::month(Date),
                                   sep = ".")) -> dat1

left_join(dat1, MEI.v2.long) %>%
  left_join(MEI.v2.lag1yr.long)%>%
  left_join(MEI.v2.lag6mo.long) %>%
  left_join(MEI.v2.byYr) %>%
  left_join(MEI.v2.lag1yr.byYr) %>%
  left_join(MEI.v2.lag6mo.byYr) %>%
  left_join(PDO.long) %>%
  left_join(PDO.lag1yr) %>%
  left_join(PDO.lag6mo) %>%
  left_join(PDO.byYr) %>%
  left_join(PDO.lag1yr.byYr) %>%
  left_join(PDO.lag6mo.byYr) %>%
  left_join(ONI.long) %>%
  left_join(ONI.lag1yr) %>%
  left_join(ONI.lag6mo) %>%
  left_join(ONI.byYr) %>%
  left_join(ONI.lag1yr.byYr) %>%
  left_join(ONI.lag6mo.byYr) -> dat1.oce

```


How does the MEI values play a role in strandings?

```{r}
p12 <- ggplot(data = dat1.oce) + 
  geom_point(aes(x = Date, 
                 y = MEI.v2, 
                 color = Species),
             size = 3)

if (save.fig)
  ggsave(plot = p12,
         device = "png",
         dpi = 600,
         filename = "figures/date_MEIv2.png")

p12
```

So, the strandings during 2015 and early 2016 might have been affected by positive MEI values - possibly in the northern region?  Next plot adds latitude to this plot and size information is dropped.

```{r}
p13 <- ggplot(data = dat1.oce) + 
  geom_point(aes(x = Date, 
                 y = MEI.v2, 
                 color = Species, 
                 size = Latitude))

if (save.fig)
  ggsave(plot = p13,
         device = "png",
         dpi = 600,
         filename = "figures/date_MEIv2_latitude.png")

p13

```

What if we lag MEI by 6 months or 1 year?

```{r}

p14 <- ggplot(data = dat1.oce) + 
  geom_point(aes(x = Date, 
                 y = MEI.v2.lag1yr, 
                 color = Species, 
                 size = Latitude))

if (save.fig)
  ggsave(plot = p14,
         device = "png",
         dpi = 600,
         filename = "figures/date_MEIv2lag1yr_latitude.png")

p14
```

```{r}

p15 <- ggplot(data = dat1.oce) + 
  geom_point(aes(x = Date, 
                 y = MEI.v2.lag6mo, 
                 color = Species, 
                 size = Latitude))

if (save.fig)
  ggsave(plot = p15,
         device = "png",
         dpi = 600,
         filename = "figures/date_MEIv2lag6mo_latitude.png")

p15
```

How did the MEI change over time?

```{r}
ggplot() + 
  geom_point(data = filter(MEI.v2.long, Year > 2008),
             aes(x = year.month.num,
                 y = MEI.v2)) + 
  geom_point(data = dat1.oce,
             aes(x = year.month.num,
                 y = 0,
                 color = Species),
             size = 3)
```

# strandings per year? 

```{r}
dat1.oce %>% 
  group_by(Year) %>%
  summarise(mean.MEI = mean(mean.MEI),
         n = n()) -> dat1.MEI.byYr

ggplot(dat1.MEI.byYr) + 
  geom_point(aes(x = mean.MEI, y = n, color = Year), size = 3)
```


```{r}
dat1.oce %>% 
  group_by(Year) %>%
  summarise(mean.MEI.lag1yr = mean(mean.MEI.lag1yr),
         n = n()) -> dat1.MEI.lag1yr.byYr

ggplot(dat1.MEI.lag1yr.byYr) + 
  geom_point(aes(x = mean.MEI.lag1yr, y = n, color = Year), size = 3)
```


```{r}
dat1.oce %>%
  group_by(Year) %>%
  summarise(mean.MEI.lag6mo = mean(mean.MEI.lag6mo),
         n = n()) -> dat1.MEI.lag6mo.byYr

ggplot(dat1.MEI.lag6mo.byYr) + 
  geom_point(aes(x = mean.MEI.lag6mo, y = n, color = Year), size = 3)
```

What if we just looked at one species?

```{r}
dat1.oce %>% filter(Species == "LO") %>% 
  group_by(Year) %>%
  summarise(mean.MEI.lag6mo = mean(mean.MEI.lag6mo),
            mean.MEI.lag1yr = mean(mean.MEI.lag1yr),
            n = n()) -> dat1.MEI.LO.lag

ggplot(dat1.MEI.LO.lag) + 
  geom_point(aes(x = mean.MEI.lag6mo, y = n, color = Year), size = 3)

ggplot(dat1.MEI.LO.lag) + 
  geom_point(aes(x = mean.MEI.lag1yr, y = n, color = Year), size = 3)

```


```{r}
dat1.oce %>% filter(Species == "LO") %>% 
  group_by(Year) %>%
  summarise(mean.PDO.lag6mo = mean(mean.PDO.lag6mo),
            mean.PDO.lag1yr = mean(mean.PDO.lag1yr),
            n = n()) -> dat1.PDO.LO.lag

ggplot(dat1.PDO.LO.lag) + 
  geom_point(aes(x = mean.PDO.lag6mo, y = n, color = Year), size = 3)

ggplot(dat1.PDO.LO.lag) + 
  geom_point(aes(x = mean.PDO.lag1yr, y = n, color = Year), size = 3)

```




```{r}
dat1.oce %>% filter(Species == "LO") %>% 
  group_by(Year) %>%
  summarise(mean.ONI.lag6mo = mean(mean.ONI.lag6mo),
            mean.ONI.lag1yr = mean(mean.ONI.lag1yr),
            n = n()) -> dat1.ONI.LO.lag

ggplot(dat1.ONI.LO.lag) + 
  geom_point(aes(x = mean.ONI.lag6mo, y = n, color = Year), size = 3)

ggplot(dat1.ONI.LO.lag) + 
  geom_point(aes(x = mean.ONI.lag1yr, y = n, color = Year), size = 3)

```

put together all potential oceanographic variables here for each year - thinking about some sort of regression analysis for the number of stranding per year as the response. 
```{r}
dat1.oce %>% #mutate(f.Year = factor(Year)) %>%
  group_by(Year, Species) %>%
  summarise(n = n()) %>%
  complete(Species, fill = list(n = 0)) %>%
  left_join(MEI.v2.byYr) %>%
  left_join(MEI.v2.lag1yr.byYr) %>%
  left_join(MEI.v2.lag6mo.byYr) %>%
  left_join(ONI.byYr) %>%
  left_join(ONI.lag1yr.byYr) %>%
  left_join(ONI.lag6mo.byYr) %>%
  left_join(PDO.byYr) %>% 
  left_join(PDO.lag1yr.byYr) %>% 
  left_join(PDO.lag6mo.byYr) -> dat1.oce.sp


```

Quick regression
Look at the collinearity among various oceanographic indices first.

```{r}
# from here: http://www.sthda.com/english/wiki/ggplot2-quick-correlation-matrix-heatmap-r-software-and-data-visualization
cormat <- round(cor(select(ungroup(dat1.oce.sp), -c(Year, Species, n))), 2)
cormat[lower.tri(cormat)] <- NA
melted_cormat <- melt(cormat)

ggplot(data = melted_cormat, 
       aes(x = Var1, y = Var2, fill = value)) + 
  geom_tile(color = "white")+
  scale_fill_gradient2(low = "blue", high = "red", mid = "white",
                       midpoint = 0, limit = c(-1,1), space = "Lab",
                       name = "Pearson\nCorrelation") + 
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1)) + 
  labs(x = "", y = "")
```
Kinda difficult to see them all... separate them out by index

```{r}
cormat_PDO <- round(cor(select(ungroup(dat1.oce.sp), contains("PDO"))), 2)
cormat_PDO[lower.tri(cormat_PDO)] <- NA

melted_cormat_PDO <- melt(cormat_PDO)

ggplot(data = melted_cormat_PDO, 
       aes(x = Var1, y = Var2, fill = value)) + 
  geom_tile(color = "white")+
  scale_fill_gradient2(low = "blue", high = "red", mid = "white",
                       midpoint = 0, limit = c(-1,1), space = "Lab",
                       name = "Pearson\nCorrelation") + 
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1)) + 
  labs(x = "", y = "")

```

mean.PDO is correlated with pretty much all except SD.PDO, SD.PDO.lag6mo, and SD.PDO.lag1yr. I'll use mean.PDO, and three other SD measures (.PDO, .lag6mo, and .lag1yr).

```{r}
cormat_ONI <- round(cor(select(ungroup(dat1.oce.sp), contains("ONI"))), 2)
cormat_ONI[lower.tri(cormat_ONI)] <- NA

melted_cormat_ONI <- melt(cormat_ONI)

ggplot(data = melted_cormat_ONI, 
       aes(x = Var1, y = Var2, fill = value)) + 
  geom_tile(color = "white")+
  scale_fill_gradient2(low = "blue", high = "red", mid = "white",
                       midpoint = 0, limit = c(-1,1), space = "Lab",
                       name = "Pearson\nCorrelation") + 
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1)) + 
  labs(x = "", y = "")
```

mean.ONI is not correlated with any other except possibly min.ONI and max.ONI.  So, I'll take mean.ONI, mean.ONI.lag6mo, and SD.ONI.lag1yr.


```{r}
cormat_MEI <- round(cor(select(ungroup(dat1.oce.sp), contains("MEI"))), 2)
cormat_MEI[lower.tri(cormat_MEI)] <- NA

melted_cormat_MEI <- melt(cormat_MEI)

ggplot(data = melted_cormat_MEI, 
       aes(x = Var1, y = Var2, fill = value)) + 
  geom_tile(color = "white")+
  scale_fill_gradient2(low = "blue", high = "red", mid = "white",
                       midpoint = 0, limit = c(-1,1), space = "Lab",
                       name = "Pearson\nCorrelation") + 
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1)) + 
  labs(x = "", y = "")
```

MEI is similar to ONI; mean.MEI is somewhat correalted with min.MEI, max.MEI, min.MEI.lag6mo, and max.MEI.lag6mo. I select mean.MEI, max.MEI.lag1yr, and SD.MEI.lag6mo.

See how indices correlate now.
```{r}
cormat_oce <- round(cor(select(ungroup(dat1.oce.sp), c("mean.PDO", 
                                                       "SD.PDO", 
                                                       "SD.PDO.lag6mo", 
                                                       "SD.PDO.lag1yr",
                                                       "mean.ONI", 
                                                       "mean.ONI.lag6mo",  
                                                       "SD.ONI.lag1yr",
                                                       "mean.MEI", "max.MEI.lag1yr", 
                                                       "SD.MEI.lag6mo"))))
cormat_oce[lower.tri(cormat_oce)] <- NA

melted_cormat_oce <- melt(cormat_oce)

ggplot(data = melted_cormat_oce, 
       aes(x = Var1, y = Var2, fill = value)) + 
  geom_tile(color = "white")+
  scale_fill_gradient2(low = "blue", high = "red", mid = "white",
                       midpoint = 0, limit = c(-1,1), space = "Lab",
                       name = "Pearson\nCorrelation") + 
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1)) + 
  labs(x = "", y = "")

```

Strong correlations for mean.PDO and others, mean.MEI and others, SD.MEI.lag6mo and others.  Drop those... 

```{r}
cormat_oce <- round(cor(select(ungroup(dat1.oce.sp), c("SD.PDO", 
                                                       "SD.PDO.lag6mo", 
                                                       "SD.PDO.lag1yr",
                                                       "mean.ONI", 
                                                       "mean.ONI.lag6mo",  
                                                       "SD.ONI.lag1yr",
                                                       "max.MEI.lag1yr"))))
cormat_oce[lower.tri(cormat_oce)] <- NA

melted_cormat_oce <- melt(cormat_oce)

ggplot(data = melted_cormat_oce, 
       aes(x = Var1, y = Var2, fill = value)) + 
  geom_tile(color = "white")+
  scale_fill_gradient2(low = "blue", high = "red", mid = "white",
                       midpoint = 0, limit = c(-1,1), space = "Lab",
                       name = "Pearson\nCorrelation") + 
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1)) + 
  labs(x = "", y = "")
```

SD.ONI.lag1yr is strongly correlated (negative and positive, respectively) with SD.PDO.lag6mo and SD.PDO.lag1yr/max.MEI.lag1yr, so remove it.  

```{r}
cormat_oce <- round(cor(select(ungroup(dat1.oce.sp), c("SD.PDO", 
                                                       "SD.PDO.lag6mo", 
                                                       "SD.PDO.lag1yr",
                                                       "mean.ONI", 
                                                       "mean.ONI.lag6mo",
                                                       "max.MEI.lag1yr"))))
cormat_oce[lower.tri(cormat_oce)] <- NA

melted_cormat_oce <- melt(cormat_oce)

ggplot(data = melted_cormat_oce, 
       aes(x = Var1, y = Var2, fill = value)) + 
  geom_tile(color = "white")+
  scale_fill_gradient2(low = "blue", high = "red", mid = "white",
                       midpoint = 0, limit = c(-1,1), space = "Lab",
                       name = "Pearson\nCorrelation") + 
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1)) + 
  labs(x = "", y = "")
```

They now all look good.  Use these variables to look at the number of strandings per year. 

```{r}
dat1.oce.sp %>% select(n, Year, Species, SD.PDO, SD.PDO.lag6mo, SD.PDO.lag1yr, 
                       mean.ONI, mean.ONI.lag6mo, max.MEI.lag1yr) -> dat1.oce.sp.1

library(lme4)

# there are only three Cc strandings... not useful? 
fit.1 <- glmer(n ~ SD.PDO + SD.PDO.lag6mo + SD.PDO.lag1yr + mean.ONI + mean.ONI.lag6mo + max.MEI.lag1yr + (1 | Species), 
             family = poisson, data = dat1.oce.sp.1)

summary(fit.1)
```

Because there were only three strandings of Cc, I remove them from the data

```{r}
dat1.oce.sp.2 <- filter(dat1.oce.sp.1, Species != "CC")

fit.2 <- glmer(n ~ SD.PDO + SD.PDO.lag6mo + SD.PDO.lag1yr + mean.ONI + mean.ONI.lag6mo + max.MEI.lag1yr + (1 | Species), 
             family = poisson, data = dat1.oce.sp.2)

summary(fit.2)

```

Only mean.ONI seems to be a somewhat useful variable, which positively affected the strandings - then again, the sample size is quite small... 

```{r}
ggplot(data = dat1.oce.sp.2) + 
  geom_point(aes(x = mean.ONI, y = n, color = Species),
             size = 3)
```



Going back to this idea of how oceanographic seasons may be affecting the number of strandings... I need to define "seasons," which may or may not correspond to year/month.  

ONI:
```{r}
ONI %>%  
  mutate(year.month = paste(YR, MON, sep = ".")) %>%
  transmute(ONI.anom = ANOM,
            year.month = year.month,
            Year = YR,
            Date = as.Date(paste(YR, MON, 15, sep = "-")))-> ONI.1

dat1 %>% left_join(select(ONI.1, c(ONI.anom, year.month)), 
                   by = "year.month") -> dat1.1

ggplot() +
  geom_path(data = filter(ONI.1, Year > 2008),
            aes(x = Date, y = ONI.anom)) + 
  geom_point(data = dat1.1,
             aes(x = Date, y = ONI.anom, color = Species),
             size = 3)

```


MEI:
```{r}
MEI.v2.long <- mutate(MEI.v2.long, 
                      Date = as.Date(paste(Year, month, 15, sep = "-")))

dat1.1 %>% left_join(select(MEI.v2.long, c(MEI.v2, year.month)), 
                   by = "year.month") -> dat1.1

ggplot() +
  geom_path(data = filter(MEI.v2.long, Year > 2008),
            aes(x = Date, y = MEI.v2)) + 
  geom_point(data = dat1.1,
             aes(x = Date, y = MEI.v2, color = Species),
             size = 3)

```



PDO:
```{r}
PDO.long %>%  
  transmute(PDO.anom = PDO,
            year.month = year.month,
            Year = YEAR,
            Date = as.Date(paste(YEAR, month, 15, sep = "-")))-> PDO.1

dat1.1 %>% left_join(select(PDO.1, c(PDO.anom, year.month)), 
                   by = "year.month") -> dat1.1

ggplot() +
  geom_path(data = filter(PDO.1, Year > 2008),
            aes(x = Date, y = PDO.anom)) + 
  geom_point(data = dat1.1,
             aes(x = Date, y = PDO.anom, color = Species),
             size = 3)

```

plot them all in one figure
```{r}
ggplot() +
  geom_path(data = filter(PDO.1, Year > 2008),
            aes(x = Date, y = PDO.anom),
            color = "red",
            size = 2) + 
  geom_path(data = filter(MEI.v2.long, Year > 2008),
            aes(x = Date, y = MEI.v2),
            color = "blue",
            size = 2) + 
  geom_path(data = filter(ONI.1, Year > 2008),
            aes(x = Date, y = ONI.anom),
            color = "yellow",
            size = 2)
#  geom_vline(data = dat1.1,
#             aes(xintercept = Date, color = Species),
#             size = 1)
```

For the most part, MEI and ONI are similar (yellow and blue) but PDO is a bit different, which makes sense.  MEI and ONI both measure likelihood of El Nino, whereas PDO is a bit more larger scale. So, makes sense to look at one of MEI/ONI and PDO.  

I wonder if these three indices may somehow linearly explain the strandings? 